Should I AirBNB My Property?

### 2017-01-31 abbDataPrep_Script.Rmd
Reorder, day 1

---
title: "AirBNB Data Preparation and Imputation"
output: html_notebook
---

@@

This file details the initial data preparation that we have applied to the AirBNB nightly data from www.AirDNA.co as well as to the long-term rental data from Australia Property Monitors (APM). This is not a 'data cleaning' exercise as we are not removing data points based on observed values.

@@

### Setup

We begin by loading the necessary libraries, set the path to the data and loading a set of custom data munging functions.

```{r message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE}

 ## Load libraries

  library(spdep)
  library(maptools)
  library(gstat)
  library(ggplot2)
  library(ggmap)
  library(xtable)
  library(chron)
  library(RColorBrewer)
  library(plyr)
  library(stringr)

 ## Set location

  # Get computer names
  comp.name <- Sys.info()['nodename']

  # Assign path based on computer name
  if(comp.name == '7020D-121777-W' | 
      comp.name == 'DESKTOP-1D7JO4J'){

      data.path <- 'c:/dropbox/research/airBNB/data/'

@@

      data.path <- 'other path'

@@

 ## Load sources

  source("c:/code/research/airbnbMelbourne/dataPrep_Functions.R")

@@

## Short Term (Airbnb) Data

We begin the data preparation with the Airbnb data.

### Load data

Next we load the raw property level and nightly level AirBNB reservation data received from AirDNA.

```{r message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, cache=TRUE}

 ## Property Information

  prop.data <- read.csv(paste0(data.path, '/raw/melbproperties.csv'), header=T)

  ## Daily rental information

  daily.data  <- read.csv(paste0(data.path, '/raw/melbDaily.csv'), header=T)

@@

### Data Transformations

To make future coding easier, we convert all field names to lower case.

```{r message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, cache=TRUE}

  names(prop.data) <- tolower(names(prop.data))
  names(daily.data) <- tolower(names(daily.data))

@@

We also then convert date values into correct R formatted dates. We also create a year created variable for the property data.

```{r message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, cache=TRUE}

 ## Daily data

  daily.data$date <- as.Date(daily.data$date)

 ## Property level data

  prop.data$created.date <- as.Date(prop.data$created.date)
  prop.data$created.year <- as.numeric(substr(prop.data$created.date, 1, 4))

@@

### Add Summary Information

We now add summary information regarding the number of bookings, number of free and blocked nights to the property level data.  This information will assist in future imputation exercises and other analyses.

We start by splitting the daily date into a list where each object is a data.frame of daily observations for a particular property.

```{r message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, cache=TRUE}

  daily.prop <- split(daily.data, daily.data$property.id)

@@

Next, we apply the **abbAnalyzeDaily()** function to each data.frame of daily observations. This process produces a list of results, by property.id.  These lists are then collapsed into a single data.frame.

```{r message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, cache=TRUE}

  daily.summ.list <- lapply(daily.prop, abbAnalyzeDaily)
  daily.summ <- rbind.fill(daily.summ.list)

@@

Finally, we add the summarized data to the existing property-level information.

```{r message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, cache=TRUE}

  prop.data <- merge(prop.data, daily.summ, by='property.id')

@@

Finally, we write the amended daily and property data to .csv files for future analysis.

```{r message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, cache=TRUE}

 write.csv(all.daily, paste0(data.path, 'prepared/dailydata.csv'), row.names=FALSE)

 write.csv(prop.data, paste0(data.path, 'prepared/propdata.csv'), row.names=FALSE)

@@

>>>>

---
title: "Data Preparation, Integration and Cleaning"
output:
  html_document: default
  html_notebook: default
---

@@

This file details the initial data preparation and cleaning that we have applied to the AirBNB nightly data from www.AirDNA.co as well as to the long-term rental data from Australia Property Monitors (APM).  Both datasets have had some initial pre-processing done, the details of which are found in **abb_str_PreProcess.R** and **abb_ltr_PreProcess.R**.

@@

### Preliminary Commands

&nbsp;

We begin with a number of preliminary commands.  First we load all of the necessary R libraries for this analysis.

```{r load_libraries, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  library(spdep)
  library(maptools)
  library(gstat)
  library(ggplot2)
  library(geosphere)
  library(ggmap)
  library(xtable)
  library(chron)
  library(plyr)
  library(sp)
  library(rgeos)
  library(reshape2)
  library(stringr)
  library(RColorBrewer)
  library(Hmisc)

We then set the paths to the data and the code, conditional on which of the authors is running the analysis.

```{r set_paths, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  # Get computer names

  comp.name <- Sys.info()['nodename']

  # Assign path based on computer name

  if(comp.name == '7020D-121777-W' | comp.name == 'DESKTOP-1D7JO4J'){

    data.path <- 'c:/dropbox/research/airBNB/data/'
    code.path <- 'c:/code/'

@@

    data.path <- 'alternate data path'
    code.path <- 'alternate data path'

@@

@@

Next, we source the necessary custom code files for the analysis.

```{r source_files, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

 # Current Project files

  source(paste0(code.path, "research/AirBNBMelbourne/abb_Functions.R"))

 # Other custom files

  source(paste0('https://raw.githubusercontent.com/andykrause/dataVizTools',
                '/master/ggPlotTools.R'))
  source(paste0('https://raw.githubusercontent.com/andykrause/datamgmttools',
                '/master/dataMungeTools.R'))
  source(paste0('https://raw.githubusercontent.com/andykrause/spatialTools',
                '/master/spatialLogitModels.R'))
  source(paste0('https://raw.githubusercontent.com/andykrause/spatialTools',
                '/master/point2surface_Tools.R'))

@@

We then set the constants used in this study:

1. **Exchange rate**: The Airbnb rates are provided in US dollars.  We convert these to Australian dollars based on the average exchange rate over the Sep 1 2015 to Aug 30 2015 period, or about 1.32 Australian dollars to 1 US dollar.

```{r set_constants, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  exch.rate <- 1.32

@@

## Loading data

&nbsp;

Next we load in the raw data.  The short-term rental (Airbnb) data is found in two separate files, one providing information on the property itself and one providing a daily record of bookings.  
```{r load_str_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Property Information

  str.data <- read.csv(paste0(data.path, '/prepared/propdata.csv'), header=T)

  # Daily ltral information

  daily.data  <- read.csv(paste0(data.path, '/prepared/dailydata.csv'), header=T)

@@

We then load the long term rental data.  This includes property level data (including the final lease rate) and listing-level data containing information on all changes to listing prices and dates, etc.

```{r load_ltr_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 # Property level long term ltral information

  ltr.data <- read.csv(paste0(data.path, '/prepared/ltpropdata.csv'), header=T)

 # Long term ltral listing information

  listing.data <- read.csv(paste0(data.path, '/prepared/ltlistdata.csv'), header=T)

@@

And, finally, we load three GIS shapefiles.  The first two denote suburb and Statistical Area 1 (sa1) boundaries in the Melbourne area. We also load a shapefile that contains the location of the beach areas along Port Phillip Bay.^[Only those east of the mouth of Yarra River as those to the west are primarily industrial land.]  After we load each we convert the names of the data fields to lower case for easier use in future coding.

```{r load_sp_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Suburb shapefile

  suburbs.shp <- readShapePoly(paste0(data.path, 'geographic/melbSuburbs.shp'),
                               proj4string=CRS("+init=epsg:4283"),
                               delete_null_obj=TRUE)
  names(suburbs.shp@data) <- tolower(names(suburbs.shp@data))

  # SA1 shapefile

  sa1s.shp <- readShapePoly(paste0(data.path, 'geographic/sa1s.shp'),
                            proj4string=CRS("+init=epsg:4283"),
                            delete_null_obj=TRUE)
  names(sa1s.shp@data) <- tolower(names(sa1s.shp@data))

  # Beach file

  beach.shp <- readShapePoly(paste0(data.path, 'geographic/portPhillipBeach.shp'),
                             proj4string=CRS("+init=epsg:4283"),
                             delete_null_obj=TRUE)
  names(beach.shp@data) <- tolower(names(beach.shp@data))

@@

## Data Preparation

&nbsp;

In this section we remove, convert and/or transform a number of data fields.  We also remove observations if they are missing critical data or fall outside of the scope of our research question.

&nbsp;

### Trim and Standardize Fields

The data contain many fields that are either not fully observed and/or are not relevant to our research question.  Additionally, the field naming conventions differ between the two datasets. We first remove unnecessary or unobserved fields and then employ a systematic field naming functions across all datasets.

We start by trimming the property-level short term data.  No fields need to be removed from the daily data.

```{r trim_st, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str.keep.flds <- c('property.id', 'host.id', 'listing.title', 'property.type',
                     'listing.type', 'created.date', 'latitude', 'longitude',
                     'last.scraped.date', 'average.daily.rate', 'overall.rating',
                     'max.guests', 'cancellation.policy', 'cleaning.fee',
                     'published.nightly.rate', 'published.weekly.rate', 
                     'published.monthly.rate', 'minimum.stay',
                     'bedrooms', 'bathrooms', 'min.date', 'max.date', 'total.days',
                     'blocked.days', 'available.days', 'created.year')

  str.data <- str.data[, str.keep.flds]

@@

We then limit the fields of the long-term property-specific data

```{r trim_lt, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  ltr.keep.flds <- c('id.key', 'geographicalid', 'eventid', 'addressid', 'activityid',
                      'flatnumber', 'streetnumber', 'streetname', 'streettype', 
                      'suburb', 'postcode', 'property_latitude',
                      'property_longitude', 'street_centroid_latitude',
                      'street_centroid_longitude', 'eventdate', 'eventprice',
                      'propertytype', 'areasize', 'bedrooms', 'baths', 'parking',
                      'hasstudy', 'hascourtyard', 'hasbalcony', 'hasairconditioning',
                      'hasgarage', 'dom')

  ltr.data <- ltr.data[,ltr.keep.flds]

@@

Next we standardize the field names.  Here we follow the short-term data convention of placing a period between words in the name.  We also shorten a number of the fields to make future coding easier.

For the short term data, it already uses the *word-.-word* convention.  Here our renaming is done to shorten certain field names. No changes are made to the names of the Airbnb daily data.

```{r rename_st, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  names(str.data) <- c('property.id', 'host.id', 'listing.title', 'property.type',
                       'listing.type', 'created.date', 'latitude', 'longitude',
                       'last.scraped.date', 'average.daily.rate', 'overall.rating',
                       'max.guests', 'cancel.policy', 'cleaning.fee',
                       'nightly.rate', 'weekly.rate', 'monthly.rate', 'min.stay',
                       'bedrooms', 'baths', 'min.date', 'max.date', 'total.days',
                       'blocked.days', 'available.days', 'created.year')

Periods are inserted between relevant words for the ltral data.

```{r rename_lt, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  names(ltr.data) <- c('id.key', 'geographical.id', 
                        'event.id', 'address.id', 'activity.id',
                        'flat.number', 'street.number', 'street.name', 'street.type',
                        'suburb', 'postcode', 'latitude', 'longitude',
                        'street.latitude', 'street.longitude', 'event.date',
                        'event.price', 'property.type', 'area.size', 'bedrooms',
                        'baths', 'parking', 'has.study', 'has.courtyard',
                        'has.balcony', 'has.aircond', 'has.garage', 'dom')

  names(listing.data) <- c('id.key', 'event.id', 'activity.id', 'address.id', 
                           'of.record', 'event.date', 'event.price', 'first.adv.date', 
                           'first.adv.price', 'last.adv.date', 'last.adv.price',
                           'event.type')

&nbsp;

### 2017-01-31 abbReturns_Script.Rmd
Reorder, day 1

### Preliminary Commands

We begin with a number of preliminary commands.  First we load all of the necessary R libraries for this analysis.

```{r load_libraries, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  library(spdep)
  library(maptools)
  library(gstat)
  library(ggplot2)
  library(geosphere)
  library(ggmap)
  library(xtable)
  library(chron)
  library(plyr)
  library(sp)
  library(rgeos)
  library(reshape2)
  library(stringr)
  library(RColorBrewer)
  library(Hmisc)

We then set the paths to the data and the code, conditional on which of the authors is running the analysis.

```{r set_paths, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  # Get computer names

  comp.name <- Sys.info()['nodename']

  # Assign path based on computer name

  if(comp.name == '7020D-121777-W' | comp.name == 'DESKTOP-1D7JO4J'){

    data.path <- 'c:/dropbox/research/airBNB/data/'
    code.path <- 'c:/code/'

  } else {

    data.path <- 'alternate data path'
    code.path <- 'alternate data path'

Next, we source the necessary custom code files for the analysis.

```{r source_files, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  source(paste0(code.path, "dataviztools/ggplottools.R"))
  source(paste0(code.path, "datamgmttools/dataMungetools.R"))
  source(paste0(code.path, "research/AirBNBMelbourne/analysis_Functions.R"))
  source(paste0(code.path, "research/AirBNBMelbourne/dataPrep_Functions.R"))
  source("c:/code/spatialtools/point2surface_Tools.R")
  source("c:/code/spatialtools/spatiallogitmodels.R")

We then set the constants used in this study:

1. **Exchange rate**: The Airbnb rates are provided in US dollars.  We convert these to Australian dollars based on the average exchange rate over the Sep 1 2015 to Aug 30 2015 period, or about 1.32 Australian dollars to 1 US dollar.

```{r set_constants, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  exch.rate <- 1.32

@@

### Data Pre-processing

Both the AirBNB and the long term rental data require some measure of data pre-processing and due to the run times of these processes they are done outside of this document.  For the AirBNB data, the pre-processing is primarily concerned with imputing missing daily observations for a small number of properties in the dataset.  Full information on this process can be found in the **airbnbDataClean.Rmd** and **dataCleanFunctions.R** files.  Do note that most of the imputed data is ignored by our choice of study time period.

The long term data received from the data provider is both wide and long and very inefficiently stored.  To aid in computation time, the pre-processing involves the trimming of fields, labeling of observation types and reconfiguration of data and can be found in the **fixAPMData.R** script.

## Loading data

Next we load in the raw data.  The Airbnb data is found in two separate files, one providing information on the property itself and one providing a daily record of bookings.  
```{r load_abb_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Property Information

  abb.data <- read.csv(paste0(data.path, '/prepared/propdata.csv'), header=T)

  # Daily ltral information

  daily.data  <- read.csv(paste0(data.path, '/prepared/dailydata.csv'), header=T)

We then load the long term rental data.  This includes property level data (including the final lease rate) and listing-level data containing information on all changes to listing prices and dates, etc.

```{r load_ltr_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 # Property level long term ltral information

  ltr.data <- read.csv(paste0(data.path, '/prepared/ltpropdata.csv'), header=T)

 # Long term ltral listing information

  listing.data <- read.csv(paste0(data.path, '/prepared/ltlistdata.csv'), header=T)

And, finally, we load three GIS shapefiles.  The first two denote suburb and Statistical Area 1 (sa1) boundaries in the Melbourne area. We also load a shapefile that contains the location of the beach areas along Port Phillip Bay.^[Only those east of the mouth of Yarra River as those to the west are primarily industrial land.]

```{r load_sp_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Suburb shapefile

  # SA1 shapefile

  sa1s.shp <- readShapePoly(paste0(data.path, 'geographic/sa1s.shp'),
                            proj4string=CRS("+init=epsg:4283"),
                            delete_null_obj=TRUE)

  # Beach file

  beach.shp <- readShapePoly(paste0(data.path, 'geographic/portPhillipBeach.shp'),
                             proj4string=CRS("+init=epsg:4283"),
                             delete_null_obj=TRUE)

## Data Preparation

In this section we remove, convert and/or transform a number of data fields.  We also remove observations if they are missing critical data or fall outside of the scope of our research question.

### Trim and Standardize Fields

We start by converting all field names to lower case to simplify future coding.  This is important as R is case sensitive.

```{r lower_names, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  names(abb.data) <- tolower(names(abb.data))
  names(daily.data) <- tolower(names(daily.data))

  names(ltr.data) <- tolower(names(ltr.data))
  names(listing.data) <- tolower(names(listing.data))

 # Geographic data

  names(sa1s.shp@data) <- tolower(names(sa1s.shp@data))
  names(suburbs.shp@data) <- tolower(names(suburbs.shp@data))

The data contain many fields that are either not fully observed and/or are not relevant to our research question.  Additionally, the field naming conventions differ between the two datasets.  We first remove unnecessary or unobserved fields and then employ a systematic field naming functions across all datasets.

We start by trimming the property-level short term data.  No fields need to be removed from the daily data.

```{r trim_st, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  abb.keep.flds <- c('property.id', 'host.id', 'listing.title', 'property.type',
                     'listing.type', 'created.date', 'latitude', 'longitude',
                     'last.scraped.date', 'average.daily.rate', 'overall.rating',
                     'max.guests', 'cancellation.policy', 'cleaning.fee',
                     'published.nightly.rate', 'published.weekly.rate', 
                     'published.monthly.rate', 'minimum.stay',
                     'bedrooms', 'bathrooms', 'min.date', 'max.date', 'total.days',
                     'blocked.days', 'available.days', 'created.year')

  abb.data <- abb.data[, abb.keep.flds]

We then limit the fields of the long-term property-specific data

```{r trim_lt, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  ltr.keep.flds <- c('id.key', 'geographicalid', 'eventid', 'addressid', 'activityid',
                      'flatnumber', 'streetnumber', 'streetname', 'streettype', 
                      'suburb', 'postcode', 'property_latitude',
                      'property_longitude', 'street_centroid_latitude',
                      'street_centroid_longitude', 'eventdate', 'eventprice',
                      'propertytype', 'areasize', 'bedrooms', 'baths', 'parking',
                      'hasstudy', 'hascourtyard', 'hasbalcony', 'hasairconditioning',
                      'hasgarage', 'dom')

  ltr.data <- ltr.data[,ltr.keep.flds]

Next we standardize the field names.  Here we follow the short-term data convention of placing a period between words in the name.  We also shorten a number of the fields to make future coding easier.

For the short term data, it already uses the *word-.-word* convention.  Here our renaming is done to shorten certain field names. No changes are made to the names of the Airbnb daily data.

```{r rename_st, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  names(abb.data) <- c('property.id', 'host.id', 'listing.title', 'property.type',
                       'listing.type', 'created.date', 'latitude', 'longitude',
                       'last.scraped.date', 'average.daily.rate', 'overall.rating',
                       'max.guests', 'cancel.policy', 'cleaning.fee',
                       'nightly.rate', 'weekly.rate', 'monthly.rate', 'min.stay',
                       'bedrooms', 'baths', 'min.date', 'max.date', 'total.days',
                       'blocked.days', 'available.days', 'created.year')

Periods are inserted between relevant words for the ltral data.

```{r rename_lt, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  names(ltr.data) <- c('id.key', 'geographical.id', 
                        'event.id', 'address.id', 'activity.id',
                        'flat.number', 'street.number', 'street.name', 'street.type',
                        'suburb', 'postcode', 'latitude', 'longitude',
                        'street.latitude', 'street.longitude', 'event.date',
                        'event.price', 'property.type', 'area.size', 'bedrooms',
                        'baths', 'parking', 'has.study', 'has.courtyard',
                        'has.balcony', 'has.aircond', 'has.garage', 'dom')

  names(listing.data) <- c('id.key', 'event.id', 'activity.id', 'address.id', 
                           'of.record', 'event.date', 'event.price', 'first.adv.date', 
                           'first.adv.price', 'last.adv.date', 'last.adv.price',
                           'event.type')

@@

The dates in the long term data are not standardized.  To fix this issue we create a small function that will remove the unnecessary digits and spaces and then convert to a standard date form.  We then apply this formula to the three date fields in the long term data (both listing and rental).  We also create a year field for the event date.

```{r fix_lt_time, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 # Build custom function

  fixLTDates <- function(x){
    temp.date <- as.character(x)
    temp.date <- str_replace_all(temp.date, ' 0:00', '')
    temp.date <- str_replace_all(temp.date, '/20', '/')
    return(as.Date(temp.date, "%d/%m/%y"))   
  }

  # Convert event date

  ltr.data$event.date <- fixLTDates(ltr.data$event.date)

  # Extract Year of event date
  ltr.data$event.year <- as.numeric(substr(ltr.data$event.date, 1, 4))

  # Converted listing data date fields

  listing.data$first.adv.date <- fixLTDates(listing.data$first.adv.date)
  listing.data$last.adv.date <- fixLTDates(listing.data$last.adv.date)

>>>>

@@

<!-- &nbsp; -->

<!-- ### Data Pre-processing -->

<!-- &nbsp; -->

<!-- Both the AirBNB and the long term rental data require some measure of data pre-processing and due to the run times of these processes they are done outside of this document.  For the AirBNB data, the pre-processing is primarily concerned with imputing missing daily observations for a small number of properties in the dataset.  Full information on this process can be found in the **airbnbDataClean.Rmd** and **dataCleanFunctions.R** files.  Do note that most of the imputed data is ignored by our choice of study time period. -->

<!-- The long term data received from the data provider is both wide and long and very inefficiently stored.  To aid in computation time, the pre-processing involves the trimming of fields, labeling of observation types and reconfiguration of data and can be found in the **fixAPMData.R** script.  -->

<!-- &nbsp; -->

<!-- ## Loading data -->

<!-- &nbsp; -->

<!-- Next we load in the raw data.  The Airbnb data is found in two separate files, one providing information on the property itself and one providing a daily record of bookings.   -->
<!-- ```{r load_abb_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!--   # Property Information -->

<!--   abb.data <- read.csv(paste0(data.path, '/prepared/propdata.csv'), header=T) -->

<!--   # Daily ltral information -->

<!--   daily.data  <- read.csv(paste0(data.path, '/prepared/dailydata.csv'), header=T) -->

<!-- ``` -->

<!-- We then load the long term rental data.  This includes property level data (including the final lease rate) and listing-level data containing information on all changes to listing prices and dates, etc.  -->

<!-- ```{r load_ltr_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!--  # Property level long term ltral information -->

<!--   ltr.data <- read.csv(paste0(data.path, '/prepared/ltpropdata.csv'), header=T) -->

<!--  # Long term ltral listing information -->

<!--   listing.data <- read.csv(paste0(data.path, '/prepared/ltlistdata.csv'), header=T) -->

<!-- ``` -->

<!-- And, finally, we load three GIS shapefiles.  The first two denote suburb and Statistical Area 1 (sa1) boundaries in the Melbourne area. We also load a shapefile that contains the location of the beach areas along Port Phillip Bay.^[Only those east of the mouth of Yarra River as those to the west are primarily industrial land.] -->

<!-- ```{r load_sp_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!--   # Suburb shapefile -->

<!--   suburbs.shp <- readShapePoly(paste0(data.path, 'geographic/melbSuburbs.shp'), -->
<!--                                proj4string=CRS("+init=epsg:4283"), -->
<!--                                delete_null_obj=TRUE) -->

<!--   # SA1 shapefile -->

<!--   sa1s.shp <- readShapePoly(paste0(data.path, 'geographic/sa1s.shp'), -->
<!--                             proj4string=CRS("+init=epsg:4283"), -->
<!--                             delete_null_obj=TRUE) -->

<!--   # Beach file -->

<!--   beach.shp <- readShapePoly(paste0(data.path, 'geographic/portPhillipBeach.shp'), -->
<!--                              proj4string=CRS("+init=epsg:4283"), -->
<!--                              delete_null_obj=TRUE) -->

<!-- ``` -->

<!-- &nbsp; -->

<!-- ## Data Preparation -->

<!-- &nbsp; -->

<!-- In this section we remove, convert and/or transform a number of data fields.  We also remove observations if they are missing critical data or fall outside of the scope of our research question.  -->

<!-- &nbsp; -->

<!-- ### Trim and Standardize Fields -->

<!-- &nbsp; -->

<!-- We start by converting all field names to lower case to simplify future coding.  This is important as R is case sensitive.   -->

<!-- ```{r lower_names, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!--  # Geographic data   -->

<!--   names(sa1s.shp@data) <- tolower(names(sa1s.shp@data)) -->
<!--   names(suburbs.shp@data) <- tolower(names(suburbs.shp@data)) -->

<!-- ``` -->

<!-- The data contain many fields that are either not fully observed and/or are not relevant to our research question.  Additionally, the field naming conventions differ between the two datasets.  We first remove unnecessary or unobserved fields and then employ a systematic field naming functions across all datasets. -->

<!-- We start by trimming the property-level short term data.  No fields need to be removed from the daily data.  -->

<!-- ```{r trim_st, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!--   abb.keep.flds <- c('property.id', 'host.id', 'listing.title', 'property.type', -->
<!--                      'listing.type', 'created.date', 'latitude', 'longitude', -->
<!--                      'last.scraped.date', 'average.daily.rate', 'overall.rating', -->
<!--                      'max.guests', 'cancellation.policy', 'cleaning.fee', -->
<!--                      'published.nightly.rate', 'published.weekly.rate',  -->
<!--                      'published.monthly.rate', 'minimum.stay', -->
<!--                      'bedrooms', 'bathrooms', 'min.date', 'max.date', 'total.days', -->
<!--                      'blocked.days', 'available.days', 'created.year')  -->

<!--   abb.data <- abb.data[, abb.keep.flds] -->

<!-- ``` -->

<!-- We then limit the fields of the long-term property-specific data -->

<!-- ```{r trim_lt, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!--   ltr.keep.flds <- c('id.key', 'geographicalid', 'eventid', 'addressid', 'activityid', -->
<!--                       'flatnumber', 'streetnumber', 'streetname', 'streettype',  -->
<!--                       'suburb', 'postcode', 'property_latitude', -->
<!--                       'property_longitude', 'street_centroid_latitude', -->
<!--                       'street_centroid_longitude', 'eventdate', 'eventprice', -->
<!--                       'propertytype', 'areasize', 'bedrooms', 'baths', 'parking', -->
<!--                       'hasstudy', 'hascourtyard', 'hasbalcony', 'hasairconditioning', -->
<!--                       'hasgarage', 'dom')  -->

<!--   ltr.data <- ltr.data[,ltr.keep.flds] -->

<!-- ``` -->

<!-- Next we standardize the field names.  Here we follow the short-term data convention of placing a period between words in the name.  We also shorten a number of the fields to make future coding easier.   -->

<!-- For the short term data, it already uses the *word-.-word* convention.  Here our renaming is done to shorten certain field names. No changes are made to the names of the Airbnb daily data.  -->

<!-- ```{r rename_st, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!--   names(abb.data) <- c('property.id', 'host.id', 'listing.title', 'property.type', -->
<!--                        'listing.type', 'created.date', 'latitude', 'longitude', -->
<!--                        'last.scraped.date', 'average.daily.rate', 'overall.rating', -->
<!--                        'max.guests', 'cancel.policy', 'cleaning.fee', -->
<!--                        'nightly.rate', 'weekly.rate', 'monthly.rate', 'min.stay', -->
<!--                        'bedrooms', 'baths', 'min.date', 'max.date', 'total.days', -->
<!--                        'blocked.days', 'available.days', 'created.year') -->

<!-- ``` -->

<!-- Periods are inserted between relevant words for the ltral data.  -->

<!-- ```{r rename_lt, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!--   names(ltr.data) <- c('id.key', 'geographical.id',  -->
<!--                         'event.id', 'address.id', 'activity.id', -->
<!--                         'flat.number', 'street.number', 'street.name', 'street.type', -->
<!--                         'suburb', 'postcode', 'latitude', 'longitude', -->
<!--                         'street.latitude', 'street.longitude', 'event.date', -->
<!--                         'event.price', 'property.type', 'area.size', 'bedrooms', -->
<!--                         'baths', 'parking', 'has.study', 'has.courtyard', -->
<!--                         'has.balcony', 'has.aircond', 'has.garage', 'dom') -->

<!--   names(listing.data) <- c('id.key', 'event.id', 'activity.id', 'address.id',  -->
<!--                            'of.record', 'event.date', 'event.price', 'first.adv.date',  -->
<!--                            'first.adv.price', 'last.adv.date', 'last.adv.price', -->
<!--                            'event.type') -->

<!-- ``` -->

<!-- &nbsp; -->

@@

<!-- The dates in the long term data are not standardized.  To fix this issue we create a small function that will remove the unnecessary digits and spaces and then convert to a standard date form.  We then apply this formula to the three date fields in the long term data (both listing and rental).  We also create a year field for the event date.  -->

<!-- ```{r fix_lt_time, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} -->

<!--  # Build custom function   -->

<!--   fixLTDates <- function(x){ -->
<!--     temp.date <- as.character(x) -->
<!--     temp.date <- str_replace_all(temp.date, ' 0:00', '') -->
<!--     temp.date <- str_replace_all(temp.date, '/20', '/') -->
<!--     return(as.Date(temp.date, "%d/%m/%y"))    -->
<!--   } -->

<!--   # Convert event date   -->

<!--   ltr.data$event.date <- fixLTDates(ltr.data$event.date) -->

<!--   # Extract Year of event date -->
<!--   ltr.data$event.year <- as.numeric(substr(ltr.data$event.date, 1, 4)) -->

<!--   # Converted listing data date fields   -->

<!--   listing.data$first.adv.date <- fixLTDates(listing.data$first.adv.date) -->
<!--   listing.data$last.adv.date <- fixLTDates(listing.data$last.adv.date) -->

<!-- ``` -->
